# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
library(plyr)
data<-read.csv("activity.csv")
data$dateTime <-strptime(paste(data$date,'00:00:00'),'%Y-%m-%d %H:%M:%S') + data$interval * 5
data$wday<-weekdays(data$dateTime)
totalStepsByDay <-ddply(data,~date,summarise,steps=sum(steps, na.rm=T))
totalStepsByFiveMinutes <-ddply(data,~interval,summarise,steps=sum(steps,na.rm=T))
totalStepsByFiveMinutes$interval <- as.numeric(row.names(totalStepsByFiveMinutes))

#data$date <- as.Date(data$date)
#data$interval<-factor(data$interval)
```


## What is mean total number of steps taken per day?
```{r}
hist(totalStepsByDay$steps,xlab="Total steps by day", main="Distribution of Daily steps taken")
xbar <- round(mean(totalStepsByDay$steps,na.rm=T),2)
x2 <- median(totalStepsByDay$steps,na.rm=T)
```
  The mean number of steps per day is **`r xbar`** and the median is **`r x2`**.
  
## What is the average daily activity pattern?
```{r}
plot(totalStepsByFiveMinutes,type="l")
maxSteps <- totalStepsByFiveMinutes[totalStepsByFiveMinutes$steps==max(totalStepsByFiveMinutes$steps),]$interval
maxStepsHours<-maxSteps*5/60

```
The five-minute interval with maximum number of steps is **`r maxSteps`th** minute (**`r maxStepsHours` hour**).


## Imputing missing values
```{r}
totalMissing<-sum(is.na(data$steps))
# use mean by interval for missing values
data$stepsImputed<-merge(data,totalStepsByFiveMinutes, by="interval",all.x=T)$steps.y
data[!is.na(data$steps),]$stepsImputed<-data[!is.na(data$steps),]$steps
# use mean by day for the rest of the missing hours
data[is.na(data$stepsImputed),]$stepsImputed<-merge(data,totalStepsByDay, by="date",all.x=T)[is.na(data$stepsImputed),]$steps.y
```
The total number of missing values is **`r totalMissing`**.
```{r}
totalStepsByDayImputed <-ddply(data,~date,summarise,steps=sum(stepsImputed))
hist(totalStepsByDayImputed$steps,xlab="Total steps by day", main="Distribution of Daily steps taken")
xbar2 <- round(mean(totalStepsByDayImputed$steps),2)
x4 <- median(totalStepsByDayImputed$steps)
```
The imputed mean is **`r xbar2`** and median is **`r x4`**, while the mean with missing values removed was **`r xbar`** and median was **`r x2`**. The histogram suggests there are some **outliers** so if we filter for those with less than 100K steps we would see the real distribution.
```{r}
hist(totalStepsByDayImputed[totalStepsByDayImputed$steps<100000,]$steps,xlab="Total steps by day", main="Distribution of Daily steps taken")
```
The difference in mean is `r xbar - xbar2` and for median is `r x2-x4`.
## Are there differences in activity patterns between weekdays and weekends?
```{r}
data$dayType<-gsub("Saturday|Sunday","weekend",data$wday)
data$dayType<-factor(gsub("Monday|Tuesday|Wednesday|Thursday|Friday","weekday",data$dayType))
totalStepsByDayType <-ddply(data,interval~dayType,summarise,steps=sum(stepsImputed))
library(lattice) 
xyplot(totalStepsByDayType$steps~totalStepsByDayType$interval|totalStepsByDayType$dayType,
       type="l",
       main="Distribution of Daily steps taken",
       ylab="Number of Steps",  
        xlab="Five minute interval",layout=c(1,2))
```

