# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
source("load_data.R")

dataset <- load_data()
# raw_dataset is the dataset without conversion of datatypes
raw_dataset <- dataset

# need to convert steps to numeric after loading
dataset$steps <- as.numeric(dataset$steps)
```

## What is mean total number of steps taken per day?
```{r}
source("mean_total_steps_every_day.R")

total_steps_per_day <- get_total_steps_per_day(dataset)

  ## draw the histogram
  hist(total_steps_per_day$total_steps, col = "red", xlab = "Number of steps", main = "Total Number of Steps Taken Each Day")

mean_total_steps <- mean(total_steps_per_day$total_steps)
median_total_steps <- median(total_steps_per_day$total_steps)
```

**Mean** total number of steps taken per day is: **`r format(mean_total_steps, scientific=FALSE, digits=4) `**

**Median** total number of steps taken per day is: **`r format(median_total_steps, scientific=FALSE, digits=4) `**

## What is the average daily activity pattern?
```{r}
source("time_series.R")

mean_steps_per_interval <- get_mean_steps_per_interval(dataset)

  # Create the time series plot
  library(ggplot2)
  ggplot(mean_steps_per_interval, aes(x=intervals, y=mean_steps)) + 
    geom_line(colour="#0000CC") +
    xlab("Intervals") +
    ylab("Mean steps") +
    ggtitle("Mean steps across all days for all 5-min intervals")

max_interval <- find_max_interval(mean_steps_per_interval)
```
**Interval** with the **maximum** mean steps is: **`r format(max_interval[1, "intervals"], scientific=FALSE, digits=4) `**

It has recorded the following mean steps : **`r format(max_interval[1, "max_mean"], scientific=FALSE, digits=4) `**

## Imputing missing values
```{r}
source("imputing_missing_values.R")

missing_values <- find_count_missing_values(raw_dataset)
```
Number of **missing** values :**`r format(missing_values[1, "missing"], scientific=FALSE, digits=4) `**

Now replacing NAs with the mean steps in all the 5-min intervals
```{r}

replaced_with_5min_mean_dataset <- replace_missing_values(dataset, mean_steps_per_interval)

```

Now creating the histogram for the new dataset with the values replaced
```{r}
new_total_steps_per_day <- get_total_steps_per_day(replaced_with_5min_mean_dataset)

  ## draw the histogram
  hist(new_total_steps_per_day$total_steps, col = "green", xlab = "Number of steps", main = "Total steps every day where NAs replaced by 5 min interval means")

new_mean_total_steps <- mean(new_total_steps_per_day$total_steps)
new_median_total_steps <- median(new_total_steps_per_day$total_steps)
```
**Mean** total number of steps taken per day for dataset with NAs replaced is: **`r format(new_mean_total_steps, scientific=FALSE, digits=4) `**

**Median** total number of steps taken per day for dataset with NAs replaced is: **`r format(new_median_total_steps, scientific=FALSE, digits=4) `**

Do these values differ from the estimates from the first part of the assignment? 

**Yes they differ from the first part.**

What is the impact of imputing missing data on the estimates of the total daily number of steps?

**The mean and median have increased after the imputting. Also the mean and median have now converged to be the same value.**

## Are there differences in activity patterns between weekdays and weekends?
```{r}
source("weekday_weekend.R")
# determine day of the week for each record
dataset_with_days <- add_weekday_weekend(dataset)
# have two subsets: 1 for weekdays 1 for weekends
weekday_subset <- get_weekday_records(dataset_with_days)
weekend_subset <- get_weekend_records(dataset_with_days)
# apply the get_mean_steps_per_interval for each subset
mean_steps_per_interval_weekday <- get_mean_steps_per_interval(weekday_subset)
mean_steps_per_interval_weekend <- get_mean_steps_per_interval(weekend_subset)
# add the column weekday and assign the value "weekday" for the weekday subset after getting the mean_steps_per_interval
mean_steps_per_interval_weekday$weekday <- c("weekday")
# add the column weekday and assign the value "weekend" for the weekend subset after getting the mean_steps_per_interval
mean_steps_per_interval_weekend$weekday <- c("weekend")

# merge the two subsets vertically
mean_steps_per_interval_by_weekday <- rbind(mean_steps_per_interval_weekend, mean_steps_per_interval_weekday)

  # Create the time series plot
  library(ggplot2)
  ggplot(mean_steps_per_interval_by_weekday, aes(x=intervals, y=mean_steps,group=weekday)) + 
    geom_line(colour="#0000CC") +
    xlab("Interval") +
    ylab("Number of steps") +
    facet_wrap( ~ weekday, nrow = 2) +
    ggtitle("Mean steps across weekdays / weekends for all 5-min intervals")

```

**YES, there are differences in activity patterns between weekdays and weekends.**